---
title: "Exploration of the iris data frame"
author: "Ana Cecilia Leon Morales"
output:
    html_document:
        theme: cosmo
        toc: true
        keep_md: true
---

#Familiarizing with the data frame

##See the entire data frame

As the data set *iris* belongs to the `datasets` package, which is one of the default packages on R, the only thing we should do to see its data is typing the next line on console: 
```R
iris
```

The result looks like this:
```{r, echo=FALSE}
iris
```

##The head and tail commands

As we could notice, is very uncomfortable and unhelpful to see the entire data frame. So, we can use some commands to visualize a part of the data frame. For example:

1. The **head** command will return the first six rows in the data frame.

```{r}
head(iris)
```

**Note:** We can also obtain the first $n$ rows in the data frame by putting the parameter $n$ inside function head. 

```{r}
#This line will return the first n = 10 rows in the data frame iris
head(iris,n=10)
```

2. The **tail** command will return by dfault the last six rows of the data frame. Furthermore, in this command the parameter $n$ works in the same way as with the head command:

```{r}
#This line will return the last n = 8 rows in the data frame iris
tail(iris,n=8)
```

This first exploration allowed us to notice that this data frame has five different variables, which refers to: 

**Hint**: See documentation by typyng `help(iris)` on the console to know more details about this data frame content.

| No. | Variable's name | Description                                 |
|-----|-----------------|---------------------------------------------|
|  1  | Sepal.Length    | The length of flowers' sepal in centimeters |
|  2  | Sepal.Width     | The width of flowers' sepal in centimeters  |
|  3  | Petal.Length    | The length of flowers' petal in centimeters  |
|  4  | Petal.Width     | The width of flowers' petal in centimeters  |
|  5  | Species         | The flowers' specie                         |

# What is the structure of the iris data frame?

The command `str(iris)` can help us to know th structure of the data frame. This is, t know the number of observations and variables, the type of data which is saved in every variables, and its values or levels.

For instance, the structure of the data frame is the following:

```{r,echo=FALSE}
str(iris)
```

# Summary measurements
As this is the first look to the date frame, it is useful to obtain certain statistical measuments such as the **mean**, **median**, **standar deviation**, etc.

## Statistical functions

We can obtain this measurements for an specific variable in the data frame by tiyping the following commands:

```r
mean(iris$Petal.Length)
median(iris$Petal.Length)
sd(iris$Petal.Length)
```
which returnd the results:

```{r,echo=FALSE}
mean(iris$Petal.Length)
median(iris$Petal.Length)
sd(iris$Petal.Length)
```

That means the petal length's mean of the flowers, which is calculated by the following formula

$$\frac{1}{150} \sum{Petal.Length}$$
will be equal to $3.758$. Moreover, the median of will be $4.35$ and the standard deviation to $1.765$.

## Summary function

It is possible to calculate other statistical measurements in all data frame by using the **summary** function:

```{r}
summary(iris)
```

## Some descriptive plots

In order to visualize the behaviour of the numeric variables on data set, we can generate the following graph:


```{r,echo=FALSE}
plot(iris[,1:4],col=topo.colors(150))
```

Finally, we can visualize the counts of our categorical variable by:
```{r,echo=FALSE}
b<-barplot(table(iris$Species),col="skyblue",border="white",ylim=c(0,60))
text(b, table(iris$Species), labels = table(iris$Species), pos =3)
```








